---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# strata

<!-- badges: start -->
[![R-CMD-check](https://github.com/asenetcky/strata/actions/workflows/R-CMD-check.yaml/badge.svg)](https://github.com/asenetcky/strata/actions/workflows/R-CMD-check.yaml)
[![Codecov test coverage](https://codecov.io/gh/asenetcky/strata/graph/badge.svg)](https://app.codecov.io/gh/asenetcky/strata)
[![CRAN status](https://www.r-pkg.org/badges/version/strata)](https://CRAN.R-project.org/package=strata)
<!-- badges: end -->

The goal of strata is to provide a framework for workflow automation
and reproducible analyses for R users and teams who may not have access
to many modern automation tooling and/or are otherwise 
resource-constrained. Strata aims to be simple and allow users to
adopt it with minimal changes to existing code and use whatever automation
they have access to.  There is only one target file users will need to automate, `main.R`, which will run through the entire project with the 
settings they specified as they build their strata of code.  Strata is 
designed to get out of the users' way and play nice with packages like
`renv`, `cronR`, and `taskscheduleR`.

## ⚙️ Installation

To install the latest CRAN release, just run:

```{r eval=FALSE}
install.packages("strata")
```


You can install the development version of strata from [GitHub](https://github.com/) with:

``` r
# install.packages("pak")
pak::pak("asenetcky/strata")
```

## ✨ Features

- Easily turn your R project into an automated one OR a one-click affair
- Easily view your project's execution plan - no more hunting and pecking
around to find out which script sources the next
- No more juggling multiple automated tasks - there is only a single automation
target - `main.R` or `main()` function
- Simple and consistent built-in logging
- Manage code execution in `.toml` files
- Quick build options for rapid prototyping

## 🎯 Target Audience

`strata` is primarily for developers who are looking for simplicity in
a framework they can adopt in their R projects. `strata` will help users quickly and consistently execute their code.  Configurations are stored in 
plain text .toml files that can be tracked with version control. Users who aren't looking to learn a new syntax, or change the way they write their code may feel right at home because `strata` only requires users to build out the file structure a specific way using the provided functions.  Users will
be able to easily 'one-click' execute their entire project in the sequence
they specify, and/or easily implement a layer of automation down the road.

Developers with limited access to advanced automation tools will find `strata`
to be a great way to prep their projects to implement the task scheduling 
automation they already have on their computers.  Users with access to more 
advanced tooling or looking for more detailed features may not have their needs
met by `strata` but may still appreciate the simplicity and logging for 
rapid prototyping.
 

## 🚀 Getting Started

`strata` provides users with framework for easier 
automation with the tools they already have at hand.
Users will want to make a folder for their `strata`
project. `strata` works all by itself but shines best when 
bundled inside of an RStudio project folder and 
coupled with the [`renv`](https://cran.r-project.org/package=renv) package.


### Build Strata

After loading and attaching the `strata` package users will want to 
start hollowing out spaces for their code to live.  Calling
`build_stratum()` and providing a name and path
to your project folder will add a 'stratum' to project, as well
as a `main.R` script and a `.strata.toml` file. 

```{r, project-stratum}
library(strata)

# Make a folder for your strata project
my_project_folder <- fs::dir_create(fs::file_temp())

# build_stratum creates a folder, a stratum,  where you can group together
# similar code into sub-folder/s called a lamina/ae

# pro tip: build_stratum invisibly returns the stratum path
stratum_path <-
  build_stratum(
    stratum_name = "project_setup",
    project_path = my_project_folder,
    order = 1
  )

# let's take a look at what was made

fs::dir_tree(my_project_folder, recurse = TRUE, all = TRUE)

# let's take a look at that .toml file
view_toml(fs::path(my_project_folder, "strata", ".strata.toml"))

# our stratum is empty, let's change that in the next section
```


### Build Laminae

Next users will want to call `build_lamina()` with a name
and path to your stratum you created in the previous step. 
This creates a sub-folder of your stratum where you R code will
live, as well as a `.laminae.toml`.  It's good to group 
like-code together inside of a 'lamina'. Users can have as many
stratum as needed with as many laminae and their associated R 
scripts that users deem necessary.  


```{r, laminae}
# remember we still have that stratum path from the previous section
# let's build some laminae you might see inside a stratum called 'project_setup'
# remember you can 1 or more R script inside of a lamina

# This could be a lamina that contains code that
# sets up your libraries for the project
build_lamina(
  stratum_path = stratum_path,
  lamina_name = "libraries",
  order = 1
)

# This could be a lamina that contains code that API keys and other credentials
# and/or makes use of the excellent `keyring` package
build_lamina(
  stratum_path = stratum_path,
  lamina_name = "authentication",
  order = 2
)

# This could be a lamina that contains code for setting up your
# database connections etc...
build_lamina(
  stratum_path = stratum_path,
  lamina_name = "connections",
  order = 3
)

# Don't worry if you forget to specify an order, strata will assign one for you!
# Always check that the order assigned is the order you want

fs::dir_tree(my_project_folder, recurse = TRUE, all = TRUE)
view_toml(fs::path(stratum_path, ".laminae.toml"))
```


### Adding R Scripts

Now that users have their stratum and laminae set up, they can
begin adding their R scripts to the laminae.  Users can drag and
drop their existing code into the laminae folders or write their
code directly in the laminae.

```{r placeholder-code}
# let's fill in the laminae with some placeholder code
auth_path <- fs::path(stratum_path, "authentication", "auth_code.R")
lib_path <- fs::path(stratum_path, "libraries", "lib_code.R")
conn_path <- fs::path(stratum_path, "connections", "conn_code.R")

authentication_code <- "print('I am your authentication setup code')"
libraries_code <- "print('I am your library setup code')"
connections_code <- "print('I am your connection setup code')"

purrr::walk(
  c(auth_path, lib_path, conn_path),
  \(r_file) fs::file_create(r_file)
)

purrr::walk2(
  c(auth_path, lib_path, conn_path),
  c(authentication_code, libraries_code, connections_code),
  \(r_file, code) cat(code, file = r_file, append = TRUE)
)
```

### Taking a Look

Adding all these folders and files is well and good, but how do users keep
track of it all? `strata` has you covered. This readme has showcased a few 
already, but let's recap and then look at the entire execution plan.


```{r proj-structure}
# let's take a look at the project structure now
fs::dir_tree(my_project_folder, recurse = TRUE, all = TRUE)

# Look at all those files now - let's grab the paths of only the .tomls
survey_tomls(my_project_folder)
```

Users can now `view_toml()` to see the contents of the `.toml` files
or edit/replace their contents as well.

The observability work-horse for most users will likely be `survey_strata()`.

```{r survey-strata}
survey_strata(my_project_folder)
```

Users can run `survey_strata()` to get all the need-to-know about their project.
It is important to note that `survey_strata()` "sees" the project through the
eyes of the `.toml` files, and _not_ the actual file structure, which might
not reflect the true nature of the strata project. If a stratum or lamina wasn't
built with `build_stratum()` or `build_lamina()` or added by hand to the
`.tomls` it will not show up in the survey. This provides users with the
flexibility to weave their strata project in and out of other projects or test
out new ideas without affecting the execution plan of the project.


### Running the Project

`main.R` and its associated function `main()` is the entry point 
to your project and the target that users will automate the 
execution of. When executed `main()` will read those `.toml` files
and begin sourcing the pipelines in the order specified by the 
user/.toml files, and within a stratum it will execute the 
laminae in the order specified by the user and their specific
.toml file.  Within a lamina the scripts will be sourced 
however the user's operating system has ordered the scripts,
often alphabetically.

There are two ways for a uaer to "run" a project:

Users can source the `main.R` script that was auto-generated when
building a stratum.

```{r source-file}
source(fs::path(my_project_folder, "main.R"))
```

Users can also just execute the `main()` function directly, all that is needed
is the path to the project.  Users can also silence the logging messages if
they desire, however, any messages generated by the code in the laminae will
still be printed to the console.


```{r main-function}
main(my_project_folder, silent = TRUE)
```

That's the core functionality of `strata`! Users build their project with 
`strata` and then it stays out of their way.

## ⚡ Going a Step Further

We have our strata project, now what?  
While it might be useful to organize all the disparate parts of your project,
and to turn it into a one-click affair, we can do better.


### Implementing Automation

Users may not have access to the lastest and greatest
automation tools and orchestrators, but users likely have access to 
something like windows task scheduler, cron jobs or similar.  In fact there
are already two great packages that can help users automate their R scripts. 
- [`cronR`](https://cran.r-project.org/package=cronR) and 
- [`taskscheduleR`](https://cran.r-project.org/package=taskscheduleR).

`strata` makes it easy to automate an entire project worth of code with
those packages.  Just point those packages' functions to `main.R` or a 
script with the `main()` function and that's it.

Using cronR:

```{r example-cron, eval=FALSE}
library(cronR)

tmp <- fs::dir_create(fs::file_temp())
build_quick_strata_project(tmp, 2, 2)

## Run our project every thursday at 0700
file <- fs::path(tmp, "main.R")
cmd <- cron_rscript(file)

cmd

cron_add(cmd, frequency = "monthly", at = "7AM", days_of_week = "THU")
cron_njobs()

cron_ls()
cron_clear(ask = TRUE)

fs::dir_delete(tmp)
```

Using taskscheduleR:

```{r example-task, eval=FALSE}
library(taskscheduleR)

tmp <- fs::dir_create(fs::file_temp())
build_quick_strata_project(tmp, 2, 2)

## Run our project every thursday at 0700
taskscheduler_create(
  taskname = "strata",
  rscript = fs::path(tmp, "main.R"),
  schedule = "WEEKLY",
  starttime = "07:00",
  days = "THU",
  startdate = "12/01/2024"
)

taskscheduler_ls() |>
  dplyr::filter(TaskName == "strata") |>
  dplyr::select(HostName:Status)

taskscheduler_delete(taskname = "strata")

fs::dir_delete(tmp)
```

## 🎁 Additional Features

<details><summary> Consistent, easy to use logging </summary>

### Logging

`strata` provides basic, but consistent logging functions that
are used at run time, and are available for use inside of users'
code if they desire. 

- Logging can be routed to stdout or stderr which can then be piped to wherever
users want to store their logs.
- Logging format is consistent -
 - The timestamp is always 26 characters long, including square brackets
 - The timestamp is always YEAR-MONTH-DAY HOUR:MINUTE:SECOND.XXXX
 - The log level follows the timestamp with a space and ends with a colon
 - The message follows the log level and is separated by a space
- Logging is kept intentionally simple, but can still be parsed with 
`survey_log()` and the resulting tibble analyzed however the user wishes

```{r logging}
# The core logging function is log_message, and users can mold it to their needs
log_message(
  message = "This is a message",
  level = "INFO", # Use whatever "level" you want
  out_or_err = "OUT" # Send to stdout
)

# Users can sending warnings
log_message(
  message = "This is a warning",
  level = "WARN",
  out_or_err = "ERR" # Send to stderr if user wants
)

# log_error() is a wrapper around log_message that sends messages to stderr
log_error("This is an error message")

# log_total_time() is a simple function that  always prints the time
# difference in seconds

duration <-
  log_total_time(
    begin = Sys.time(),
    end = Sys.time() + 999
  )

log_message(
  paste("This took", duration, "seconds"),
  level = "INFO",
  out_or_err = "OUT"
)
```

`log_message()` and `log_error()` invisibly return a character string copy
of their output.

```{r}
log_output <- log_message("I am a log message")
log_output
```

`survey_log()` will return a tibble of your log file contents for you to view.
_Only_ output from the `log_message()` or `log_error()` functions will be 
returned from the log, any other output will be ignored.  In this example
we are building a fake log line by line, normally the log would be built
by having the stdout and stderr piped to a log file by the user or automation.

```{r}
# Build an example log
example_log <- fs::file_create(fs::file_temp(ext = "log"))

# show file is empty
readr::read_lines(example_log)

# add log message
line1 <- log_message("strata started")
line1 <- paste0(line1, "\n")
cat(line1, file = example_log, append = TRUE)

# nonsense output to be ignored
line2 <- "I am a line of nonsense"
line2 <- paste0(line2, "\n")
cat(line2, file = example_log, append = TRUE)

# another log message
line3 <- log_message("strata finished")
line3 <- paste0(line3, "\n")
cat(line3, file = example_log, append = TRUE)

# parse log and return tibble
# notice that the line numbers are preserved, even if content is ignored
survey_log(example_log)

# clean up
fs::file_delete(example_log)
```


</details>


<details><summary> Quick build projects </summary>

###  Quick Build

`strata` has two "quick" build options for users.

`build_quick_strata_project()` will create a project with a stratum and laminae
with some placeholder code.  This is great for testing out ideas or deploying
current code into a strata project as quickly as possible and then sorting out
the names and .tomls later.
  
```{r}
tmp <- fs::dir_create(fs::file_temp())
build_quick_strata_project(
  project_path = tmp,
  num_strata = 2,
  num_laminae_per = 3
)

fs::dir_tree(tmp)
```

`build_outline_strata_project()` will create a project with strata and laminae
based on an outline dataframe provided by the user.  This is great for users who 
have a specific vision in mind already.

```{r}
tmp <- fs::dir_create(fs::file_temp())

outline <-
  dplyr::tibble(
    project_path = tmp,
    stratum_name = c(
      rep("env_setup", 2),
      rep("data_pull", 2),
      "data_wrangle",
      "build_model",
      "build_report"
    ),
    stratum_order = c(
      rep(1, 2),
      rep(2, 2),
      3,
      4,
      5
    ),
    lamina_name = c(
      c("authentication", "connections"),
      c("sql-source1", "sql-source2"),
      "clean_data",
      "tidy_models",
      "quarto_report"
    ),
    lamina_order = c(1, 2, 1, 2, 1, 1, 1),
    skip_if_fail = FALSE
  )

dplyr::glimpse(outline)

build_outlined_strata_project(outline)

fs::dir_tree(tmp)
survey_strata(tmp)
```


</details>


<details><summary> Execute any part of your strata project ad-hoc </summary>
 
###  Ad-hoc

The ability to run the entire project with `main()` is great, but if users want
to run pieces of the project ad-hoc, it would be inconvenient to have to source
the disparate pieces of code individually, and possibly error-prone as well.
Users can execute their project piecemeal with `adhoc_stratum()` and
`adhoc_lamina()`, and `adhoc()`.

`adhoc_stratum()` will source the code inside of _every_ lamina in the stratum,
while ignoring all other strata.

```{r}
tmp <- fs::dir_create(fs::file_temp())

build_quick_strata_project(tmp, 2, 2)

adhoc_stratum(
  stratum_path = fs::path(tmp, "strata", "stratum_1"),
  silent = FALSE
)
```

`adhoc_lamina()` will execute _only_ the specified lamina and the code therein
contained, ignoring all other laminae inside the same stratum and all other
strata.

```{r}
# using our quick build project from above
adhoc_lamina(
  lamina_path = fs::path(tmp, "strata", "stratum_2", "s2_lamina_2"),
  silent = FALSE
)
```

In interactive sessions, `adhoc()` will execute a stratum or lamina based on the
name provided by the user.  Users won't have to remember the paths.  `adhoc()`
will default to the current working directory, or users can optionally provide a
project path. If there are multiple *exact* matches, `adhoc()` will prompt users
in the console to pick their intended target.  If users opt to not to be
prompted `adhoc()` will execute the first match it finds.

```{r, eval=FALSE}
# using our quick build project from above
adhoc(
  name = "s2_lamina_1",
  project_path = tmp,
  silent = FALSE,
  prompt = FALSE
)
```


</details>


<details><summary> Easy to manage configuration inside of R or out </summary>

### Managing .tomls

Users are able to find and read .toml files in their project with the 
`survey_tomls()` and `view_toml()` functions.  
Users also have options for editing their .toml files as well.
Users will always have the option to use their favorite text editor to edit
any one of the .toml files.  The files may be hidden, but `survey_tomls()`
will provide the proper paths.  Users can then confirm their edits with
`view_toml()` and confirm that their expected changes appear in the tibble.

Users can opt to work with the `edit_toml()` function and stay completely inside
of R code.  It is _strongly_ advised users save a copy of their target .toml
file in memory using `view_toml()` to fall back on.  Users can then take a 
copy of that and edit it.  All that is required of `edit_toml()` is a file path
to the original .toml file and the tibble of the content they wish to replace
the original .toml file with.

```{r edit-toml}
# create temporary folder
tmp <- fs::dir_create(fs::file_temp())

# quick build for demonstration purposes
build_quick_strata_project(tmp, 2, 2)

# survey the .tomls
toml_list <-
  survey_tomls(tmp)

toml_list


# create copies
original_strata_toml <- view_toml(toml_list[1])
original_strata_toml

original_lamina1_toml <- view_toml(toml_list[2])
original_lamina1_toml

# original execution plan
original_plan <- survey_strata(tmp)
```

The originals are all backed up, now users can edit the .toml files as they see
fit.

```{r edit-toml-strata}
# edit the strata .toml
# swap the order of the strata execution
new_strata_toml <-
  original_strata_toml |>
  dplyr::mutate(
    order = c(2, 1)
  )

# make the edits
edit_toml(
  original_toml_path = toml_list[1],
  new_toml_dataframe = new_strata_toml
)

# check the order
view_toml(toml_list[1])
```

Users will notice that in addition to the changes made, a backup of the file was
created in the same directory with the extension `.bak`.  Version control
should be the first line of defense to protect against loss, but this is a nice 
safety net.  The newest backups will always clobber the oldest, so care needs to
be taken.

```{r edit-toml-lamina}
# edit the lamina toml
# swap the order of the lamina execution
# make laminae skip if fail
new_lamina_toml <-
  original_lamina1_toml |>
  dplyr::mutate(
    order = c(2, 1),
    skip_if_fail = TRUE
  )

# make the edits
edit_toml(
  original_toml_path = toml_list[2],
  new_toml_dataframe = new_lamina_toml
)

# check the order and skip_if_fail
view_toml(toml_list[2])
```

Users can now check the execution plan to see if their changes have taken 
effect and how they have changed the entire project's execution order.

```{r execution-plan}
# check execution order for the entire project
survey_strata(tmp)
```


</details>


<details><summary> Ability to skip failing lamina </summary>

### Skip if Fail

Users can specify that a lamina should be skipped if it fails by setting the
`skip_if_fail` to `TRUE` when building the laminae, or doing the same inside of
the `.laminae.toml` files.  This is useful when users are prototyping new
additions to their strata project. Use with *caution*, this is not intended to
be a replacement for more robust error handling implemented by the user within
their own code.

```{r}
tmp <- fs::dir_create(fs::file_temp())

build_quick_strata_project(tmp, 2, 2)

# let's add a lamina that skips if it fails
build_lamina(
  stratum_path = fs::path(tmp, "strata", "stratum_1"),
  lamina_name = "bad_lamina",
  order = 2,
  skip_if_fail = TRUE
)

# lets add the code that will fail
bad_code <- "stop('I am bad code')"
bad_path <- fs::path(tmp, "strata", "stratum_1", "bad_lamina", "bad_code.R")
fs::file_create(bad_path)
cat(bad_code, file = bad_path)

# let's run the project
main(tmp)
```

Users will see in the log that the lamina that failed was skipped and the rest
of the project executed.

</details>
