---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# strata

<!-- badges: start -->
[![R-CMD-check](https://github.com/asenetcky/strata/actions/workflows/R-CMD-check.yaml/badge.svg)](https://github.com/asenetcky/strata/actions/workflows/R-CMD-check.yaml)
[![Codecov test coverage](https://codecov.io/gh/asenetcky/strata/graph/badge.svg)](https://app.codecov.io/gh/asenetcky/strata)
[![CRAN status](https://www.r-pkg.org/badges/version/strata)](https://CRAN.R-project.org/package=strata)
<!-- badges: end -->

The goal of strata is to provide a framework for workflow automation
and reproducible analyses for R users and teams who may not have access
to many modern automation tooling and/or are otherwise 
resource-constrained. Strata aims to be simple and allow users to
adopt it with minimal changes to existing code and use whatever automation
they have access to.  There is only one target file users will need to automate, `main.R`, which will run through the entire project with the 
settings they specified as they build their strata of code.  Strata is 
designed to get out of the users' way and play nice with packages like
`renv`, `cronR`, and `taskscheduleR`.

## Installation

To install the latest CRAN release, just run:

```{r eval=FALSE}
install.packages("strata")
```


You can install the development version of strata from [GitHub](https://github.com/) with:

``` r
# install.packages("pak")
pak::pak("asenetcky/strata")
```

## ✨ Features

- Turn your R project into an automated one OR a one-click affair
- Only a single automation target - `main.R` or `main()` function
- Simple and consistent built-in logging
- Manage code execution in `.toml` files
- Quick build options for rapid prototyping


## 🚀 Getting Started Using `strata`

`strata` provides users with framework for easier 
automation with the tools they already have at hand.
Users will want to make a folder for their `strata`
project. `strata` works all by itself but shines best when 
bundled inside of an RStudio project folder and 
coupled with the [`renv`](https://cran.r-project.org/package=renv) package.

After loading and attaching the `strata` package users will want to 
start hollowing out spaces for their code to live.  Calling
`build_stratum()` and providing a name and path
to your project folder will add a 'stratum' to project, as well
as a `main.R` script and a `.strata.toml` file. 

```{r, project_and_stratum}
library(strata)

# Make a folder for your strata project
my_project_folder <- fs::dir_create(fs::file_temp())

# build_stratum creates a folder, a stratum,  where you can group together
# similar code into sub-folder/s called a lamina/ae

# pro tip: build_stratum invisibly returns the stratum path
stratum_path <-
  build_stratum(
    stratum_name = "project_setup",
    project_path = my_project_folder,
    order = 1
  )

# let's take a look at what was made

fs::dir_tree(my_project_folder, recurse = TRUE, all = TRUE)

# let's take a look at that .toml file
view_toml(fs::path(my_project_folder, "strata", ".strata.toml"))

# our stratum is empty, let's change that in the next section
```

Next users will want to call `build_lamina()` with a name
and path to your stratum you created in the previous step. 
This creates a sub-folder of your stratum where you R code will
live, as well as a `.laminae.toml`.  It's good to group 
like-code together inside of a 'lamina'. Users can have as many
stratum as needed with as many laminae and their associated R 
scripts that users deem necessary.  


```{r, laminae}
# remember we still have that stratum path from the previous section
# let's build some laminae you might see inside a stratum called 'project_setup'
# remember you can 1 or more R script inside of a lamina

build_lamina(
  stratum_path = stratum_path,
  lamina_name = "libraries",
  order = 1
)

build_lamina(
  stratum_path = stratum_path,
  lamina_name = "authentication",
  order = 2
)

build_lamina(
  stratum_path = stratum_path,
  lamina_name = "connections",
  order = 3
)

# Don't worry if you forget to specify an order, strata will assign one for you!
# Always check that the order assigned is the order you want

fs::dir_tree(my_project_folder, recurse = TRUE, all = TRUE)
view_toml(fs::path(stratum_path, ".laminae.toml"))
```


`main.R` and its associated function `main` is the entry point 
to your project and the target that users will automate the 
execution of. When executed `main` will read those `.toml` files
and begin sourcing the pipelines in the order specified by the 
user/.toml files, and within a stratum it will execute the 
laminae in the order specified by the user and their specific
.toml file.  Within a lamina the scripts will be sourced 
however the user's operating system has ordered the scripts,
often alphabetically.

`strata` provides basic, but consistent logging functions that
are used at run time, and are available for use inside of users'
code if they desire.


## Examples


Here is a standard example using strata:

```{r standard-example}
library(strata)

tmp <- fs::dir_create(fs::file_temp())
strata::build_stratum(
  project_path = tmp,
  stratum_name = "first_stratum",
  order = 1
)

stratum_path <-
  fs::path(
    tmp, "strata", "first_stratum"
  )
strata::build_lamina(
  stratum_path = stratum_path,
  lamina_name = "first_lamina",
  order = 1
)
strata::build_lamina(
  stratum_path = stratum_path,
  lamina_name = "second_lamina",
  order = 2
)

lamina_path1 <- fs::path(stratum_path, "first_lamina")
lamina_path2 <- fs::path(stratum_path, "second_lamina")
code_path1 <- fs::path(lamina_path1, "my_code1.R")
code_path2 <- fs::path(lamina_path2, "my_code2.R")


my_code1 <- fs::file_create(code_path1)
my_code2 <- fs::file_create(code_path2)
cat(file = my_code1, "print('Hello, World!')")
cat(file = my_code2, "print('Goodbye, World!')")

source(fs::path(tmp, "main.R"))
fs::dir_delete(tmp)
```

Users can also opt to use quick build functions if speed is the priority
and the naming conventions are not important. 

```{r quick-build-example}
tmp <- fs::dir_create(fs::file_temp())

strata::build_quick_strata_project(
  project_path = tmp,
  num_strata = 3,
  num_laminae_per = 2
)

fs::dir_tree(tmp)
fs::dir_delete(tmp)
```

