---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# strata

<!-- badges: start -->
[![R-CMD-check](https://github.com/asenetcky/strata/actions/workflows/R-CMD-check.yaml/badge.svg)](https://github.com/asenetcky/strata/actions/workflows/R-CMD-check.yaml)
[![Codecov test coverage](https://codecov.io/gh/asenetcky/strata/graph/badge.svg)](https://app.codecov.io/gh/asenetcky/strata)
[![CRAN status](https://www.r-pkg.org/badges/version/strata)](https://CRAN.R-project.org/package=strata)
<!-- badges: end -->

The goal of strata is to provide a framework for workflow automation
and reproducible analyses for R users and teams who may not have access
to many modern automation tooling and/or are otherwise 
resource-constrained. Strata aims to be simple and allow users to
adopt it with minimal changes to existing code and use whatever automation
they have access to.  There is only one target file users will need to automate, `main.R`, which will run through the entire project with the 
settings they specified as they build their strata of code.  Strata is 
designed to get out of the users' way and play nice with packages like
`renv`, `cronR`, and `taskscheduleR`.

## Installation

To install the latest CRAN release, just run:

```{r eval=FALSE}
install.packages("strata")
```


You can install the development version of strata from [GitHub](https://github.com/) with:

``` r
# install.packages("pak")
pak::pak("asenetcky/strata")
```

## ✨ Features

- Easily turn your R project into an automated one OR a one-click affair
- Easily view your project's execution plan - no more hunting and pecking around
to find out which script sources the next
- No more juggling multiple automated tasks - there is only a single automation
target - `main.R` or `main()` function
- Simple and consistent built-in logging
- Manage code execution in `.toml` files
- Quick build options for rapid prototyping


## 🚀 Getting Started

`strata` provides users with framework for easier 
automation with the tools they already have at hand.
Users will want to make a folder for their `strata`
project. `strata` works all by itself but shines best when 
bundled inside of an RStudio project folder and 
coupled with the [`renv`](https://cran.r-project.org/package=renv) package.


### 🛠️ Build Strata

After loading and attaching the `strata` package users will want to 
start hollowing out spaces for their code to live.  Calling
`build_stratum()` and providing a name and path
to your project folder will add a 'stratum' to project, as well
as a `main.R` script and a `.strata.toml` file. 

```{r, project-stratum}
library(strata)

# Make a folder for your strata project
my_project_folder <- fs::dir_create(fs::file_temp())

# build_stratum creates a folder, a stratum,  where you can group together
# similar code into sub-folder/s called a lamina/ae

# pro tip: build_stratum invisibly returns the stratum path
stratum_path <-
  build_stratum(
    stratum_name = "project_setup",
    project_path = my_project_folder,
    order = 1
  )

# let's take a look at what was made

fs::dir_tree(my_project_folder, recurse = TRUE, all = TRUE)

# let's take a look at that .toml file
view_toml(fs::path(my_project_folder, "strata", ".strata.toml"))

# our stratum is empty, let's change that in the next section
```


### 🛠️ Build Laminae

Next users will want to call `build_lamina()` with a name
and path to your stratum you created in the previous step. 
This creates a sub-folder of your stratum where you R code will
live, as well as a `.laminae.toml`.  It's good to group 
like-code together inside of a 'lamina'. Users can have as many
stratum as needed with as many laminae and their associated R 
scripts that users deem necessary.  


```{r, laminae}
# remember we still have that stratum path from the previous section
# let's build some laminae you might see inside a stratum called 'project_setup'
# remember you can 1 or more R script inside of a lamina

# This could be a lamina that contains code that
# sets up your libraries for the project
build_lamina(
  stratum_path = stratum_path,
  lamina_name = "libraries",
  order = 1
)

# This could be a lamina that contains code that API keys and other credentials
# and/or makes use of the excellent `keyring` package
build_lamina(
  stratum_path = stratum_path,
  lamina_name = "authentication",
  order = 2
)

# This could be a lamina that contains code for setting up your
# database connections etc...
build_lamina(
  stratum_path = stratum_path,
  lamina_name = "connections",
  order = 3
)

# Don't worry if you forget to specify an order, strata will assign one for you!
# Always check that the order assigned is the order you want

fs::dir_tree(my_project_folder, recurse = TRUE, all = TRUE)
view_toml(fs::path(stratum_path, ".laminae.toml"))
```


### Adding R Scripts

Now that users have their stratum and laminae set up, they can
begin adding their R scripts to the laminae.  Users can drag and
drop their existing code into the laminae folders or write their
code directly in the laminae.

```{r placeholder-code}
# let's fill in the laminae with some placeholder code
auth_path <- fs::path(stratum_path, "authentication", "auth_code.R")
lib_path <- fs::path(stratum_path, "libraries", "lib_code.R")
conn_path <- fs::path(stratum_path, "connections", "conn_code.R")

authentication_code <- "print('I am your authentication setup code')"
libraries_code <- "print('I am your library setup code')"
connections_code <- "print('I am your connection setup code')"

purrr::walk(
  c(auth_path, lib_path, conn_path),
  \(r_file) fs::file_create(r_file)
)

purrr::walk2(
  c(auth_path, lib_path, conn_path),
  c(authentication_code, libraries_code, connections_code),
  \(r_file, code) cat(code, file = r_file, append = TRUE)
)
```

### 🔎 Taking a Look

Adding all these folders and files is well and good, but how do users keep
track of it all? `strata` has you covered. This readme has showcased a few 
already, but let's recap and then look at the entire execution plan.


```{r proj-structure}
# let's take a look at the project structure now
fs::dir_tree(my_project_folder, recurse = TRUE, all = TRUE)

# Look at all those files now - let's grab the paths of only the .tomls
survey_tomls(my_project_folder)
```

Users can now `view_toml()` to see the contents of the `.toml` files
or edit/replace their contents as well.

The observability work-horse for most users will likely be `survey_strata()`.

```{r survey-strata}
survey_strata(my_project_folder)
```

Users can run `survey_strata()` to get all the need-to-know about their project.
It is important to note that `survey_strata()` "sees" the project through the
eyes of the `.toml` files, and _not_ the actual file structure, which might
not reflect the true nature of the strata project. If a stratum or lamina wasn't
built with `build_stratum()` or `build_lamina()` or added by hand to the
`.tomls` it will not show up in the survey. This provides users with the
flexibility to weave their strata project in and out of other projects or test
out new ideas without affecting the execution plan of the project.


### Running the Project

`main.R` and its associated function `main()` is the entry point 
to your project and the target that users will automate the 
execution of. When executed `main()` will read those `.toml` files
and begin sourcing the pipelines in the order specified by the 
user/.toml files, and within a stratum it will execute the 
laminae in the order specified by the user and their specific
.toml file.  Within a lamina the scripts will be sourced 
however the user's operating system has ordered the scripts,
often alphabetically.

There are two ways to "run" your project:

Users can source the `main.R` script that was auto-generated when
building a stratum.

```{r source-file}
source(fs::path(my_project_folder, "main.R"))
```

Users can also just execute the `main()` function directly, all that is needed
is the path to the project.  Users can also silence the logging messages if
they desire, however, any messages generated by the code in the laminae will
still be printed to the console.


```{r main-function}
# or execute the main function directly
# users can silence the logging as well
# however
main(my_project_folder, silent = TRUE)
```




`strata` provides basic, but consistent logging functions that
are used at run time, and are available for use inside of users'
code if they desire.

## Going a Step Further

We have our strata project, now what?  Users can source `main.R` or any script
that calls `main()` to run the entire project.  Which can already be helpful,
but we can do better. Users may not have access to the lastest and greatest
automation tools and orchestrators, but most users should have access to 
something like windows task scheduler, cron jobs or similiar.  In fact there
are already two great packages that can help users automate their R scripts, 
[`cronR`](https://cran.r-project.org/package=cronR) and 
[`taskscheduleR`](https://cran.r-project.org/package=taskscheduleR).

Now instead of juggling what to automate and when, or having to remember to
which script sources which, users can automate the entire project with a single
task, have consistent logging messages and use `strata` to have eyes on
the entire project and its execution plan.


## Examples


Here is a standard example using strata:

```{r standard-example}
library(strata)

tmp <- fs::dir_create(fs::file_temp())
strata::build_stratum(
  project_path = tmp,
  stratum_name = "first_stratum",
  order = 1
)

stratum_path <-
  fs::path(
    tmp, "strata", "first_stratum"
  )
strata::build_lamina(
  stratum_path = stratum_path,
  lamina_name = "first_lamina",
  order = 1
)
strata::build_lamina(
  stratum_path = stratum_path,
  lamina_name = "second_lamina",
  order = 2
)

lamina_path1 <- fs::path(stratum_path, "first_lamina")
lamina_path2 <- fs::path(stratum_path, "second_lamina")
code_path1 <- fs::path(lamina_path1, "my_code1.R")
code_path2 <- fs::path(lamina_path2, "my_code2.R")


my_code1 <- fs::file_create(code_path1)
my_code2 <- fs::file_create(code_path2)
cat(file = my_code1, "print('Hello, World!')")
cat(file = my_code2, "print('Goodbye, World!')")

source(fs::path(tmp, "main.R"))
fs::dir_delete(tmp)
```

Users can also opt to use quick build functions if speed is the priority
and the naming conventions are not important. 

```{r quick-build-example}
tmp <- fs::dir_create(fs::file_temp())

strata::build_quick_strata_project(
  project_path = tmp,
  num_strata = 3,
  num_laminae_per = 2
)

fs::dir_tree(tmp)
fs::dir_delete(tmp)
```

